\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-11-30 Manuscript/manuscript.tex
PUB

@@

The first inspection that was performed on the dataset was related to the density distributions of a selection of continuous features that are known, from the celestial mechanics, to be important for the prediction of the asteroids dangerousness. These are reported in Fig. \ref{Density_relevant}. Next we moved on a more systematic analysis with FADM and mutual information. In Fig. \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig. \ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong anti-correlation of the mean motion with the semi-major axis. This is due to the Kepler law discussed in Appendix A. Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid. As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass. On the other side the fact that relative velocity is correlated with the diameter is a spurios correlation. At this point one can ask why the mean motion and the relative velocity are orthogonal: this point will be clarified from a theoretical point of view in appendix A and also with graphical models. Briefly the motion of an object on an ellipse as seen from a focus is not uniform: this is faster as the two bodies approaches. Beside this inspection a further analysis based on the concept of mutual information (summarized in the previous chapter) was performed: its result is reported in Fig. \ref{Mutual_information}. This figure summarize the ranking of the features considered in the dataset for the dangerousness classification of the asteroids according to the following expression \cite{kratzer2018varrank}

@@

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Perihelion_Distance.pdf}
    \subcaption{ \begin{center}
    a) Perihelion Distance
    \end{center}}
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Eccentricity.pdf}
    \subcaption{ \begin{center}
    b) Eccentricity
    \end{center}}
    \vspace{4ex}
  \end{minipage} \\
    \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Min_orbit_intersection.pdf}
    \subcaption{ \begin{center}
    c) Min orbit intersection
    \end{center}}
    \vspace{4ex}
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Absolute_magnitude.pdf}
    \subcaption{ \begin{center}
    d) Semi Major Axis
    \end{center}}
    \vspace{4ex}
  \end{minipage}
\caption{Comparison between the density distributions of hazardous (red) and non-hazardous (light blue) asteroids for a selected set of features that, according to the theory, are interesting.}
\label{Density_relevant}
\end{figure}

@@

\cite{murray1999solar}:

>>>>

\epigraph{
\textit{In its efforts to learn as much as possible about nature, modern physics has found that certain things can never be “known” with certainty. Much of our knowledge must always remain uncertain. The most we can know is in terms of probabilities.}\\Richard P. Feynman (1918-1988)
}

@@

The first inspection that was performed on the dataset was related to the density distributions of a selection of continuous features that are known, from the celestial mechanics, to be important for the prediction of the asteroids dangerousness. These are reported in Fig. \ref{Density_relevant}. Next we moved on a more systematic analysis with FADM and mutual information. In Fig. \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig. \ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong anti-correlation of the mean motion with the semi-major axis. This is due to the Kepler laws discussed in Appendix A. Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid. As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass. On the other side the fact that relative velocity is correlated with the diameter is a spurios correlation. At this point one can ask why the mean motion and the relative velocity are orthogonal: this point will be clarified from a theoretical point of view in appendix A and also with graphical models. Briefly the motion of an object on an ellipse as seen from a focus is not uniform: this is faster as the two bodies approaches. Beside this inspection a further analysis based on the concept of mutual information (summarized in the previous chapter) was performed: its result is reported in Fig. \ref{Mutual_information}. This figure summarize the ranking of the features considered in the dataset for the dangerousness classification of the asteroids according to the following expression \cite{kratzer2018varrank}

@@

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Perihelion_Distance.pdf}
    \subcaption{ \begin{center}
    a) Perihelion Distance
    \end{center}}
    \vspace{4ex}
  \end{minipage} 
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Eccentricity.pdf}
    \subcaption{ \begin{center}
    b) Eccentricity
    \end{center}}
    \vspace{4ex}
  \end{minipage} \\
    \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Min_orbit_intersection.pdf}
    \subcaption{ \begin{center}
    c) Min orbit intersection
    \end{center}}
    \vspace{4ex}
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Absolute_magnitude.pdf}
    \subcaption{ \begin{center}
    d) Absolute magnitude
    \end{center}}
    \vspace{4ex}
  \end{minipage}
\caption{Comparison between the density distributions of hazardous (red) and non-hazardous (light blue) asteroids for a selected set of features that, according to the theory, are interesting.}
\label{Density_relevant}
\end{figure}

@@
